ACARIS: Improving Conversational AI and Human Social Skills using User Embeddings

### 2023-05-18 paper.tex
further progress, almost established a baseline; made source data more robust

\title{ACARIS: Improving Conversational AI and Human Social Skills using User Embeddings}
\author{Simon Slamka}
\date{June 1, 2023}

@@

\section{Keywords}
ACARIS, Conversational AI, Social Skills, User Embeddings, BERT, DistilBERT, Sentiment Analysis, Intent Classification, Emotion Recognition, Interest Recognition

\section{Definitions}

@@

	\item \textbf{the Transformer} - a deep learning model introduced in 2017, used primarily in the field of NLP\cite{vaswani2017attention}
	\item \textbf{BERT} - Bidirectional Encoder Representations from Transformers - a transformer-based machine learning technique for natural language processing pre-training developed by Google\cite{devlin2019bert}

@@

\section{Motivation and Premise}

@@

\section{Hypothesis}
Given enough conversational data per person, human behavior (emotional state, intent) in text communication can be predicted with a high level of confidence ($>80\%$) due to the fact that humans are, on a fundamental level, very similar to one another.

%\begin{multicols}{2}
\section{Introduction}
Interpersonal communication has always been an integral part of human lives. With the rise of the Internet and, subsequently, social media and other forms of online text communication, the manner in which we talk has changed dramatically. This has led to a drop in social skills in humans, particularly those of the last generation. ACARIS attempts to adjust for this by providing a way to analyze the emotional state, intent, and interest of text communication parties, providing them with a way to improve their social skills, while also improving the performance of conversational AI systems, which are becoming increasingly prevalent in our society, and their ability to understand human emotions, intent, and interest is becoming more and more important, especially in AI systems that directly interact with humans, such as digital assistants, chatbots, and others.

\subsection{Literature Review and Related Work}
\subsubsection{Studied Literature}
\begin{itemize}
	\item \textbf{Machine Learning with PyTorch and Scikit-Learn}\cite{raschka2022machine} by Yuxi Liu, Vahid Mirjalili, Sebastian Raschka
	\item \textbf{Hands-on Machine Learning with Scikit-Learn, Keras, and TensorFlow}\cite{geron2022hands} by Aurélien Géron
\end{itemize}

\subsection{Initial Attempts}
Our initial attempts of implementing ACARIS included the use of SVMs, DTs, LogReg, and RFs, each paired with TF-IDF in the preprocessing stage. None of these methods have proven to be accurate enough (barely reaching 50\% accuracy) for large amounts of complex language, especially casual language that often includes slang, sarcasm, emoji, and similar elements. The primary issue was the lack of these methods' ability to capture semantical and contextual aspects of our dataset. Another issue was TF-IDF's lack of word order understanding. Knowing that the Transformer\cite{vaswani2017attention} architecture is capable of capturing these aspects due to its positional encodinds and self-attention mechanism, we decided to use it as the basis for ACARIS.

\subsection{Short Intro on Transformers}
The Transformer\cite{vaswani2017attention} is a deep learning model proposed in 2017 by a team of Google Brain researchers. It is mainly used in NLP, but it has also found its way into numerous vision projects (Latent Diffusion\cite{rombach2021highresolution}/Stable Diffusion/LLaVA\cite{liu2023visual}, to name a few).

@@

\subsection{Dataset}
To build our dataset, we collected 100 thousand messages from our Discord server, classified them into 3 classes (pos, neg, neu), and augmented them with the GoEmotions\cite{demszky2020goemotions} dataset, which contains 58 thousand messages classified into 27 classes. We squashed the 27 classes into 3 classes (pos, neg, neu) and concatenated the two datasets. To minimize error caused by input data, we removed all users with less than 25 messages, URLs, markdown code blocks, Discord mentions and custom emoji, and attachments. The resulting dataset was split into 85\% training, 10\% validation, and 5\% test sets.

\subsection{User Embeddings Initialization}
For each user $u$, we loop through all messages $m$ sent by $u$ and extract these features:
\begin{itemize}
	\item \textbf{Mean word count (per all $m$)} - \(\overline{w} = \frac{1}{N} \sum_{i=1}^{N} w_i \), where $w_i$ is the word count of $i$-th message and $N$ is the total number of messages
	\item \textbf{Vocabulary richness (per all $m$)} - \(r = \frac{U}{W} \), where $U$ is the number of unique words and $W$ is the total number of words
	\item \textbf{Mean emoji count (per all $m$)} - \(\overline{e} = \frac{1}{N} \sum_{i=1}^{N} e_i \), where $e_i$ is the emoji count of $i$-th message and $N$ is the total number of messages
	\item \textbf{Mean emoticon count (per all $m$)} - \(\overline{em} = \frac{1}{N} \sum_{i=1}^{N} em_i \), where $e_i$ is the emoticon count of $i$-th message and $N$ is the total number of messages
	\item \textbf{Mean punctuation count (per all $m$)} - \(\overline{p} = \frac{1}{N} \sum_{i=1}^{N} p_i \), where $p_i$ is the punctuation count of $i$-th message and $N$ is the total number of messages
	\item \textbf{Mean sentiment score (per all $m$)} - \(\overline{s} = \frac{1}{N} \sum_{i=1}^{N} s_i \), where $s_i$ is the sentiment score of $i$-th message and $N$ is the total number of messages
	\item \textbf{Dominant/prevalent topics (per all $m$)} - the most common topics in all messages
	\item \textbf{Mean response time (per all $m$)} - \(\overline{r} = \frac{1}{N} \sum_{i=1}^{N} r_i \), where $r_i$ is the response time of $i$-th message and $N$ is the total number of messages
	\item \textbf{Mean message count (per day)} - \(\overline{m} = \frac{1}{D} \sum_{i=1}^{D} m_i \), where $m_i$ is the message count of $i$-th day and $D$ is the total number of days
	\item \textbf{Mean links (per all $m$)} - \(\overline{l} = \frac{1}{N} \sum_{i=1}^{N} l_i \), where $l_i$ is the link count of $i$-th message and $N$ is the total number of messages
	\item \textbf{Mean markdown code snippet sections (per all $m$)} - \(\overline{c} = \frac{1}{N} \sum_{i=1}^{N} c_i \), where $c_i$ is the markdown code snippet section count of $i$-th message and $N$ is the total number of messages
	\item \textbf{Mean abbreviations and acronyms (per all $m$)} - \(\overline{a} = \frac{1}{N} \sum_{i=1}^{N} a_i \), where $a_i$ is the abbreviation and acronym count of $i$-th message and $N$ is the total number of messages
	\item \textbf{Mean hashtag count (per all $m$)} - \(\overline{h} = \frac{1}{N} \sum_{i=1}^{N} h_i \), where $h_i$ is the hashtag count of $i$-th message and $N$ is the total number of messages
\end{itemize}
Then, we concatenate all the features into a single vector $e_u$ (of size $d_e$):
\begin{equation}
\begin{aligned}
&\text{\texttt{from torch import cat}} \\
&e_u = \textnormal{\texttt{cat}}([\overline{w}, r, \overline{e}, \overline{em}, \overline{p}, \overline{s}, \overline{r}, \overline{m}, \overline{l}, \overline{c}, \overline{a}, \overline{h}], \textnormal{dim}=-1)
\end{aligned}
\end{equation}
\subsection{Concatenation of User and Word Embeddings}
Given a message $m$ with its word embedding $e_w$ (of size $d_m$) from a BERT-like pre-trained model, which, in our case, was DistilBERT\cite{sanh2020distilbert}, we concatenate the user embedding $e_u$ to the word embedding $e_w$ to create a joint representation $e_\textnormal{joint}$ (of size $d_e + d_m$):
\begin{equation}
\begin{aligned}
&\text{\texttt{from torch import cat}} \\
&e_\textnormal{joint} = \textnormal{\texttt{cat}}([e_u, e_w], \textnormal{dim}=-1)
\end{aligned}
\end{equation}

>>>>

\title{ACARIS: Improving Conversational AI and Human Social Skills using User Embeddings}
\author{Simon Slamka\\
\small\textbf{OngakkenAI}}
\date{June 1, 2023}

@@

\section{Introduction}
\subsection{Keywords}
ACARIS, Conversational AI, Social Skills, User Embeddings, BERT, DistilBERT, Sentiment Analysis, Intent Classification, Emotion Recognition, Interest Recognition

\subsection{Definitions}

@@

	\item \textbf{the Transformer} - a deep learning architecture introduced in 2017, used primarily in the field of NLP\cite{vaswani2017attention}
	\item \textbf{BERT} - Bidirectional Encoder Representations from Transformers - a transformer-based machine learning model for natural language processing pre-training developed by Google\cite{devlin2019bert}

@@

\subsection{Motivation}

@@

\subsection{Hypothesis}
Given enough conversational data per person, human behavior (emotional state, intent) in text communication can be predicted with a high level of confidence ($>80\%$) due to the fact that humans are, on a fundamental level, very similar to one another.

Additionally, we posit that person-specific performance improvements can be achieved by individualizing predictions by using a unique vector representation of a person's personality, emotional state, intent, and interest, which we call a user embedding. The postulate stems in the core concept of neural networks being universal function approximators. Therefore, in theory, a neural network should be able to learn to associate a user embedding with a person's behavior in text communication and use that information to improve its predictions.

%\begin{multicols}{2}
\subsection{Premise}
Interpersonal communication has always been an integral part of human lives and is critical from the moment we're born. With the rise of the Internet and, subsequently, social media and other forms of online text communication, the manner in which we talk has changed dramatically. This has led to a drop in social skills in humans, particularly those of the last generation. ACARIS attempts to adjust for this by providing a way to analyze the emotional state, intent, and interest of text communication parties, providing them with a way to improve their social skills, while also improving the performance of conversational AI systems, which are becoming increasingly prevalent in our society, and their ability to understand human emotions, intent, and interest is becoming more and more important, especially in AI systems that directly interact with humans, such as digital assistants, chatbots, and others.

\subsection{Literature Review and Related Work}
\subsubsection{Studied Literature}
\begin{itemize}
	\item \textbf{Machine Learning with PyTorch and Scikit-Learn}\cite{raschka2022machine} by Yuxi Liu, Vahid Mirjalili, Sebastian Raschka
	\item \textbf{Hands-on Machine Learning with Scikit-Learn, Keras, and TensorFlow}\cite{geron2022hands} by Aurélien Géron
	\item \textbf{Transformers for Natural Language Processing}\cite{rothman2022transformers} by Denis Rothman, Antonio Gulli
\end{itemize}
\subsubsection{Related Work}
\begin{itemize}
	\item \textbf{Human-AI Collaboration Enables More Empathic Conversations in Text-based Peer-to-Peer Mental Health Support}\cite{sharma2022humanai}
	\item \textbf{Context Matters: Recovering Human Semantic Structure from Machine Learning Analysis of Large-Scale Text Corpora}\cite{https://doi.org/10.1111/cogs.13085}
\end{itemize}

\subsection{Initial Attempts}
Our initial attempts of implementing ACARIS included the use of SVMs, DTs, LogReg, and RFs, each paired with TF-IDF in the preprocessing stage. None of these methods have proven to be accurate enough (barely reaching 50\% accuracy) for large amounts of complex language, especially casual language that often includes slang, sarcasm, emoji, and similar elements. The primary issue was the lack of these methods' ability to capture semantical and contextual aspects of our dataset. SVMs, DTs, LogReg, and RFs weren't built specifically for sequential data. For sequential data, RNNs and CNNs were classically used. However, due to the long-term dependency problem of RNNs, we made the decision to not use them, as we believed that they would not be able to capture the long-term dependencies in our dataset. Additionally, we weren't able to guarantee that during inference, there would be no long-term dependencies.
Another issue was TF-IDF's lack of word order understanding. Knowing that the Transformer\cite{vaswani2017attention} architecture is capable of capturing these aspects due to its positional encodinds and self-attention mechanism, we decided to use it as the basis for ACARIS.

\subsection{Short Intro on Transformers}
The Transformer is a deep learning model proposed in 2017 by a team of Google Brain/Google Research researchers. It is mainly used in NLP, but it has also found its way into numerous vision projects (Latent Diffusion\cite{rombach2021highresolution}/Stable Diffusion/LLaVA\cite{liu2023visual}, to name a few).

@@

\subsection{Dataset}
To build our dataset, we collected 100 thousand messages from our Discord server, classified them into 3 classes (pos, neg, neu), and augmented them with the GoEmotions\cite{demszky2020goemotions} dataset, which contains 58 thousand messages classified into 27 classes. We squashed the 27 classes into 3 classes (pos, neg, neu) and concatenated the two datasets. To minimize error caused by input data, we removed all users with less than 25 messages, URLs, markdown code blocks, Discord mentions and custom emoji, and attachments. The resulting dataset was split into 75\% training, 10\% validation, and 15\% test sets.

\subsection{User Embeddings Initialization}
For each user $u$, we loop through all messages $m$ sent by $u$ and extract these features:
\begin{itemize}
	\item \textbf{Mean word count (per all $m$)} - \(\overline{w} = \frac{1}{N} \sum_{i=1}^{N} w_i \), where $w_i$ is the word count of $i$-th message and $N$ is the total number of messages
	\item \textbf{Vocabulary richness (per all $m$)} - \(r = \frac{U}{W} \), where $U$ is the number of unique words and $W$ is the total number of words
	\item \textbf{Mean emoji count (per all $m$)} - \(\overline{e} = \frac{1}{N} \sum_{i=1}^{N} e_i \), where $e_i$ is the emoji count of $i$-th message and $N$ is the total number of messages
	\item \textbf{Mean emoticon count (per all $m$)} - \(\overline{em} = \frac{1}{N} \sum_{i=1}^{N} em_i \), where $e_i$ is the emoticon count of $i$-th message and $N$ is the total number of messages
	\item \textbf{Mean punctuation count (per all $m$)} - \(\overline{p} = \frac{1}{N} \sum_{i=1}^{N} p_i \), where $p_i$ is the punctuation count of $i$-th message and $N$ is the total number of messages
	\item \textbf{Mean sentiment score (per all $m$)} - \(\overline{s} = \frac{1}{N} \sum_{i=1}^{N} s_i \), where $s_i$ is the sentiment score of $i$-th message and $N$ is the total number of messages
	\item \textbf{Dominant/prevalent topics (per all $m$)} - the most common topics in all messages
	\item \textbf{Mean response time (per all $m$)} - \(\overline{r} = \frac{1}{N} \sum_{i=1}^{N} r_i \), where $r_i$ is the response time of $i$-th message and $N$ is the total number of messages
	\item \textbf{Mean message count (per day)} - \(\overline{m} = \frac{1}{D} \sum_{i=1}^{D} m_i \), where $m_i$ is the message count of $i$-th day and $D$ is the total number of days
	\item \textbf{Mean links (per all $m$)} - \(\overline{l} = \frac{1}{N} \sum_{i=1}^{N} l_i \), where $l_i$ is the link count of $i$-th message and $N$ is the total number of messages
	\item \textbf{Mean markdown code snippet sections (per all $m$)} - \(\overline{c} = \frac{1}{N} \sum_{i=1}^{N} c_i \), where $c_i$ is the markdown code snippet section count of $i$-th message and $N$ is the total number of messages
	\item \textbf{Mean abbreviations and acronyms (per all $m$)} - \(\overline{a} = \frac{1}{N} \sum_{i=1}^{N} a_i \), where $a_i$ is the abbreviation and acronym count of $i$-th message and $N$ is the total number of messages
	\item \textbf{Mean hashtag count (per all $m$)} - \(\overline{h} = \frac{1}{N} \sum_{i=1}^{N} h_i \), where $h_i$ is the hashtag count of $i$-th message and $N$ is the total number of messages
\end{itemize}
These features were selected based on what we thought would most accurately represent a user's text communication behavior. We projected that no two users would have the same or closely similar values for all of these features, which would allow us to differentiate between them.
Then, we concatenate all the features into a single vector $e_u$ (of size $d_e$):
\begin{equation}
\begin{aligned}
&\text{\texttt{from torch import cat}} \\
&e_u = \textnormal{\texttt{cat}}([\overline{w}, r, \overline{e}, \overline{em}, \overline{p}, \overline{s}, \overline{r}, \overline{m}, \overline{l}, \overline{c}, \overline{a}, \overline{h}], \textnormal{dim}=-1)
\end{aligned}
\end{equation}
\subsection{Concatenation of User and Word Embeddings}
Given a message $m$ with its word embedding $e_w$ (of size $d_m$) from a BERT-like pre-trained model, which, in our case, was DistilBERT\cite{sanh2020distilbert}, we concatenate the user embedding $e_u$ to the word embedding $e_w$ to create a joint representation $e_\textnormal{joint}$ (of size $d_e + d_m$):
\begin{equation}
\begin{aligned}
&\text{\texttt{from torch import cat}} \\
&e_\textnormal{joint} = \textnormal{\texttt{cat}}([e_u, e_w], \textnormal{dim}=-1)
\end{aligned}
\end{equation}
